---
title: "Executive Summary: An Analysis of Avocados"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Valerie Chu"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git](https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git)

:::


```{r}
#| echo: false

# load packages
library(tidyverse)
library(tidymodels)
library(here)

# handle common conflicts
tidymodels_prefer()

# load packages
load(here("results/avocado_split.rda"))
load(here("results/avocado_recipe_param.rda"))
load(here("results/avocado_recipe_tree.rda"))
load(here("results/fit_null.rda"))
load(here("results/fit_lm.rda"))
load(here("results/tuned_lasso.rda"))
load(here("results/tuned_ridge.rda"))
load(here("results/tuned_bt.rda"))
load(here("results/tuned_knn.rda"))
load(here("results/tuned_rf.rda"))
load(here("results/tbl_result.rda"))

load(here("results/avocado_recipe_param_2.rda"))
load(here("results/avocado_recipe_tree_2.rda"))
load(here("results/fit_null_2.rda"))
load(here("results/fit_lm_2.rda"))
load(here("results/tuned_lasso_2.rda"))
load(here("results/tuned_ridge_2.rda"))
load(here("results/tuned_bt_2.rda"))
load(here("results/tuned_knn_2.rda"))
load(here("results/tuned_rf_2.rda"))
load(here("results/tbl_result_2.rda"))

load(here("results/metrics_final_fit.rda"))
load(here("results/final_pred_plot.rda"))

```


```{r, echo = FALSE}
# Read in the avocado dataset
avocado <- read_csv(here("data/avocado.csv")) |> 
  janitor::clean_names()

```

## Purpose

I love avocados. Using this [Avocado Prices dataset on Kaggle](https://www.kaggle.com/datasets/neuromusic/avocado-prices) ^[Kiggins, Justin. "Avocado Prices." Kaggle, 2018, https://www.kaggle.com/datasets/neuromusic/avocado-prices.], I want to predict the average price of a single avocado (`average_price`) so that I can pinpoint and eventually move to a city with the cheapest avocados...


## Major Takeaways
@fig-final-pred-plot shows the exploration of predictions vs true values: 

```{r, echo = FALSE}
#| label: fig-final-pred-plot
#| fig-cap: "The Metrics of the Final Model to the Testing Data"

final_pred_plot

```



## Conclusion

In conclusion, the kitchen sink random forest model was the best predictive model out of all model types (Null (baseline model), OLS, Lasso, Ridge, Boosted Tree, K Nearest Neighbors, and Random Forest) and across both the kitchen sink and feature engineered recipes. 

In the future, I want to boost the predictive accuracy of this model and work with updated data to predict the average price of a single avocado based on the variables in this dataset. I've accomplished my goal of predicting the average price of a single Haas avocado sold between 2015 and 2018, and now I just need an even more accurate model and an even more updated dataset to pin down the city I want to move to in 2026, a city with the cheapest avocados…
 


