---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Valerie Chu"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git](https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git)

:::


```{r}
# load packages
library(tidyverse)
library(tidymodels)
library(here)

# handle common conflicts
tidymodels_prefer()

```


```{r}
avocado <- read_csv(here("data/avocado.csv")) |> 
  janitor::clean_names()

```

## Why this data

In California where I'm from, avocados are a big thing. There's avocado toast, avocado spread, avocado in salad, and much more. That's why I was interested in a dataset about avocados.

## Data source

This dataset contains historical data on avocado prices and sales volume in multiple US markets.

This is the link to the Avocado Prices dataset on Kaggle:
**[https://www.kaggle.com/datasets/neuromusic/avocado-prices](https://www.kaggle.com/datasets/neuromusic/avocado-prices)**

I found this dataset on Kaggle. Justin Kiggins, the owner of the dataset, writes that this data was downloaded from the Hass Avocado Board website in May of 2018 & compiled into a single CSV.

There are 18,249 observations in this dataset and 14 variables. 

These variables include `average_price`, `total_volume`, and more that I think could be super interesting to analyze and use as predictors of the prices avocados will sell at or places where avocados will be cheaper at. 

These are the definitions for some of the variables: 

- `date` - The date of the observation

- `average_price` - the average price of a single avocado

- `type` - conventional or organic

- `year` - the year

- `region` - the city or region of the observation

- `total_volume` - Total number of avocados sold

- `x4046` - Total number of avocados with PLU 4046 sold

- `x4225` - Total number of avocados with PLU 4225 sold

- `x4770` - Total number of avocados with PLU 4770 sold


## Data quality & complexity check

As mentioned above, there are 18,249 observations and 14 variables. 

There is 1 date-time object. There are 10 numerical variables. There are two categorical variables. 

The dataset is complex. 

Our check for missingness reveals that there is none in this dataset. 

```{r}
naniar::gg_miss_var(avocado)

```


## Potential data issues

I don't anticipate any potential data issues. There are no missing observations. The dataset is well-documented and well-sourced. It's also very interesting. 


## Misc

The avocado dataset is clean, organized, and has lots of data. 

I'm super excited to start working with it! 
