---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Valerie Chu"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git](https://github.com/stat301-2-2024-winter/final-project-2-valerieyichu.git)

:::


```{r}
#| echo: false

# load packages
library(tidyverse)
library(tidymodels)
library(here)

# handle common conflicts
tidymodels_prefer()

```


```{r, echo = FALSE}
# Read in the avocado dataset
avocado <- read_csv(here("data/avocado.csv")) |> 
  janitor::clean_names()

```

## An Quick Overview

I'm Californian, and avocados are a big thing. There's avocado toast, avocado spread, avocado in salad, and much more. That's why I was interested in a dataset about avocados. Hopefully, this final project will help me pinpoint and eventually move to a city with the cheapest avocados...

This dataset contains historical data on avocado prices and sales volume in multiple US markets.

This is the link to the Avocado Prices dataset on Kaggle:
**[https://www.kaggle.com/datasets/neuromusic/avocado-prices](https://www.kaggle.com/datasets/neuromusic/avocado-prices)**

This data was downloaded from the Hass Avocado Board website in May of 2018 & compiled into a single CSV. There are 18,249 observations in this dataset and 14 variables. There is 1 date-time object, 10 numerical variables, 2 categorical variables and 1 id column. There is no missingness in this dataset. 

**I want to predict the prices avocados will sell at (ie `average_price`).** So this is a **regression problem**. 

These are the definitions for some of the variables: 

- `date` - The date of the observation

- `average_price` - the average price of a single avocado

- `type` - conventional or organic

- `year` - the year

- `region` - the city or region of the observation

- `total_volume` - Total number of avocados sold

- `x4046` - Total number of avocados with PLU 4046 sold

- `x4225` - Total number of avocados with PLU 4225 sold

- `x4770` - Total number of avocados with PLU 4770 sold



## Table

::: {.callout-tip icon="false"}
## Assessment Metric

The assessment metric is the RMSE. 

:::

The table of RMSEs and standard errors created below was made using kitchen recipes. 

The baseline is the null model. 

```{r, echo = FALSE}
load(here("results/tbl_result.rda"))
tbl_result |> 
  knitr::kable()

```

The best model is the random forest model because it has the lowest RMSE. 

## EDA


## How it's going

I've split and folded the data; created a kitchen sink recipe; and fitted null, OLS, lasso, ridge, boosted tree, k nearest neighbors, and random forest models. 




